\chapter{Model Setup and Execution} \label{ch:setup}

% -------------------------------------------------------------------- %
\section{Box-model} \label{sec:boxmodel}

AtChem2 is designed to build and run atmospheric chemistry box-models
based upon the Master Chemical Mechanism (MCM,
http://mcm.leeds.ac.uk/MCM/). This page explains how to set up, compile
and run an atmospheric chemistry box-model with AtChem2. The directory
structure of AtChem2 is described {[}{[}here\textbar{}1.2 Model
Structure{]}{]}. A working knowledge of the \textbf{unix shell} and its
\href{https://swcarpentry.github.io/shell-novice/reference/}{basic
commands} is \emph{required} to use the AtChem2 model.

There are two sets of inputs to AtChem2 - the mechanism file, and
configuration files.

\subsection{Mechanism file} \label{subsec:mechanism-file}

The model requires a chemical mechanism in FACSIMILE format
(\texttt{.fac}). The \textbf{mechanism file} can be downloaded from the
MCM website using the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/extract.htt}{extraction tool} or
assembled manually. The user can modify the \texttt{.fac} file as
required with a text editor. This mechanism file is converted into a
shared library and a set of associated data files in the compilation
step below. For more information on the chemical mechanism go to:
{[}{[}2.1 Chemical Mechanism{]}{]}.

\subsection{Configuration files} \label{subsec:configuration-files}

The \textbf{model configuration} is set via a number of text files
located in the \texttt{model/configuration/} directory. The text files
can be modified with a text editor. Detailed information on the
configuration files can be found in the corresponding wiki pages:

\begin{itemize}
\item
  model and solver parameters - see {[}{[}2.2 Model Parameters{]}{]} and
  {[}{[}2.3 Solver Parameters{]}{]}.
\item
  environment variables - see {[}{[}2.4 Environment Variables{]}{]}.
\item
  photolysis rates - see {[}{[}2.5 Photolysis Rates and JFAC{]}{]}.
\item
  initial concentrations of chemical species and lists of output
  variables - see {[}{[}2.6 Config Files{]}{]}.
\end{itemize}

The model constraints - chemical species, environment variables,
photolysis rates - are located in the \texttt{model/constraints/}
directory. For more information, go to: {[}{[}2.7 Constraints{]}{]}.

\subsection{Compilation} \label{subsec:compilation}

The script \texttt{build.sh} in the \texttt{tools/} directory is used to
process the chemical mechanism file (\texttt{.fac}) and to compile the
model. The script generates one Fortran file (\texttt{mechanism.f90}),
one shared library (\texttt{mechanism.so}) and four configuration files
(\texttt{mechanism.prod}, \texttt{mechanism.reac},
\texttt{mechanism.ro2}, \texttt{mechanism.species}) in the
\texttt{model/configuration/} directory. Go to the {[}{[}chemical
mechanism page \textbar{}2.1 Chemical Mechanism{]}{]} for more
information.

The script must be run from the \emph{AtChem2 main directory} and takes
four arguments (see the \textbf{Important Note 2} at the end of this
section):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  the path to the chemical mechanism file - no default (suggested:
  \texttt{model/}).
\item
  the path to the directory for the Fortran files generated from the
  chemical mechanism - default: \texttt{model/configuration/}.
\item
  the path to the directory with the configuration files - default:
  \texttt{model/configuration/}.
\item
  the path to the directory with the MCM data files - default:
  \texttt{mcm/}.
\end{enumerate}

For example, if the \texttt{.fac} file is in the \texttt{model/}
directory:

\begin{verbatim}
./tools/build.sh model/mechanism.fac model/configuration/ model/configuration/ mcm/
\end{verbatim}

An installation of AtChem2 can have multiple \texttt{model/}
directories, which may correspond to different models or different
projects; this allows the user to run more than one model at the same
time. In the following example, there are two \texttt{model/}
directories, each with their own chemical mechanism, configuration,
constraints and output:

\begin{verbatim}
AtChem2/
        | mcm/
        | model_1/
             | configuration/
             | constraints/
             | output/
             | mechanism.fac
        | model_2/
             | configuration/
             | constraints/
             | output/
             | mechanism.fac
        | obj/
        | src/
        | tools/
        | travis/
\end{verbatim}

Each model can be built by passing the correct path to the
\texttt{build.sh} script (see the \textbf{Important Note 1} at the end
of this section). For example:

\begin{verbatim}
./tools/build.sh model_1/mechanism.fac model_1/configuration/ model_1/configuration/ mcm/
./tools/build.sh model_2/mechanism.fac model_2/configuration/ model_2/configuration/ mcm/
\end{verbatim}

Compilation is required only once for a given \texttt{.fac} file. If the
user changes the configuration files, there is no need to recompile the
model. Likewise, if the constraints files are changed, there is no need
to recompile the model. This is because the model configuration and the
model constraints are read by the executable at runtime. However, if the
user makes changes to the \texttt{.fac} file, then the shared library
\texttt{model/configuration/mechanism.so} needs to be recompiled from
the source file \texttt{model/configuration/mechanism.f90} using the
\texttt{build.sh} script.

The user may want or need to change the Fortran code
(\texttt{src/*.f90}), in which case the model needs to be recompiled: if
the \texttt{.fac} file has also been changed, use the \texttt{build.sh}
script, as explained above. Otherwise, if only the Fortran code has been
changed, executing \texttt{make} from the \emph{main directory} is
enough to recompile the model.

\subsection{Execution} \label{subsec:execution}

The compilation process creates an executable file called
\texttt{atchem2} in the \emph{main directory}. The executable file takes
seven arguments, corresponding to the directories containing the model
configuration and output:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  the path to the directory for the model output - default:
  \texttt{model/output}
\item
  the path to the directory for the model output reaction rates -
  default: \texttt{model/output/reactionRates}
\item
  the path to the directory with the configuration files - default:
  \texttt{model/configuration/}.
\item
  the path to the directory with the MCM data files - default:
  \texttt{mcm/}.
\item
  the path to the directory with the data files of constrained chemical
  species - default: \texttt{model/constraints/species/}
\item
  the path to the directory with the data files of constrained
  environment variables - default:
  \texttt{model/constraints/environment/}
\item
  the path to the directory with the data files of constrained
  photolysis rates - default: \texttt{model/constraints/photolysis/}
\end{enumerate}

The model can be run by executing the \texttt{atchem2} command from the
\emph{main directory}, in which case the executable will use the default
configuration and output directories. Otherwise, the configuration and
output directories need to be specified (see the \textbf{Important Note
2} at the end of this section).

For example, if the constraints are in the default directories (or not
used), the model can be run by executing:

\begin{verbatim}
./atchem2 model/output/ model/output/reactionRates/ model/configuration/
\end{verbatim}

In the case of multiple \texttt{model/} directories, the directories
corresponding to each model need to be passed as arguments to the
\texttt{atchem2} executable. This allows the user to run two or more
models simultaneously. For example:

\begin{verbatim}
    ./atchem2 model_1/output/ model_1/output/reactionRates/ model_1/configuration/ mcm/ model_1/constraints/species/ model_1/constraints/environment/ model_1/constraints/photolysis/
    ./atchem2 model_2/output/ model_2/output/reactionRates/ model_2/configuration/ mcm/ model_2/constraints/species/ model_2/constraints/environment/ model_2/constraints/photolysis/
\end{verbatim}

\subsubsection{Important Note 1} \label{important-note-1}

As explained above, if the chemical mechanism (\texttt{.fac}) is
changed, only the shared library needs to be recompiled. This allows the
user to have only one base executable called \texttt{atchem2} in the
\emph{main directory}: when running multiple models at the same time the
user can reuse this base executable while pointing each model to the
correct shared library and configuration files.

\subsubsection{Important Note 2} \label{important-note-2}

The arguments need to be passed to the \texttt{atchem2} executable in
the exact order, as listed above. This means that if - for example - the
third argument needs to be specified, it is also necessary to specify
the first and the second arguments, even if they have the default
values. To avoid mistakes, the user can choose to always specify all the
arguments. This behaviour also applies to the \texttt{tools/build.sh}
script used to compile the model. Future versions of AtChem2 will adopt
a simpler command-line interface.

\subsection{Output} \label{subsec:output}

The model output is saved by default in the directory
\texttt{model/output/}. The location can be modified by changing the
arguments of the \texttt{atchem2} executable (see above).

The AtChem2 output files are space-delimited text files, with a header
containing the names of the variables:

\begin{itemize}
\item
  values of environment variables and concentrations of chemical
  species: \texttt{environmentVariables.output},
  \texttt{speciesConcentrations.output}.
\item
  values of photolysis rates and related parameters:
  \texttt{photolysisRates.output},
  \texttt{photolysisRatesParameters.output}.
\item
  loss and production rates of selected species (see {[}{[}2.6 Config
  Files{]}{]}): \texttt{lossRates.output},
  \texttt{productionRates.output}.
\item
  Jacobian matrix (if requested, see {[}{[}2.2 Model Parameters{]}{]}):
  \texttt{jacobian.output}.
\item
  model diagnostic variables: \texttt{finalModelState.output},
  \texttt{initialConditionsSetting.output},
  \texttt{mainSolverParameters.output}.
\end{itemize}

In addition, the reaction rates of all the reactions in the chemical
mechanism are saved in the directory \texttt{reactionRates/}: one file
for each model step, with the filename corresponding to the time in
seconds.

\subsection{Running on HPC} \label{subsec:running-on-hpc}

Atchem2 can be set up to run on a High Performance Computing (HPC)
system. Compilation and configuration are the same as for a normal
workstation. Typically, a job scheduler is used to allocate computing
resources on an HPC system. A \textbf{submission script} is therefore
needed to submit the AtChem2 models for execution.

The format and the syntax of the submission script depend on the
specific software installed on the HPC system. For instructions on how
to prepare a submission script for AtChem2, check the local
documentation or ask the HPC system administrator.

An \emph{example} submission script for the
\href{https://en.wikipedia.org/wiki/Portable_Batch_System}{Portable
Batch System (PBS)} is shown below:

\begin{verbatim}
#PBS -o atchem2.log
#PBS -e atchem2_error.log
#PBS -N base_v1
#PBS -l walltime=15:00:00
#PBS -l vmem=10gb
#PBS -m bea
#PBS -l nodes=1:ppn=1

cd ~/AtChem2/
MODELDIR="base_model_v1"
./atchem2 $MODELDIR/model/output/ $MODELDIR/model/output/reactionRates/ $MODELDIR/model/configuration/ $MODELDIR/mcm/ $MODELDIR/model/constraints/species $MODELDIR/model/environment/species $MODELDIR/photolysis/constraints/species
\end{verbatim}

% -------------------------------------------------------------------- %
\section{Chemical Mechanism} \label{sec:mechanism}

The \textbf{chemical mechanism} is the core element of an atmospheric
chemistry box-model. In AtChem2, the mechanism file is written in
FACSIMILE format and has the extension \texttt{.fac}. The FACSIMILE
format is used to describe chemical reactions in the commercial
\href{http://www.mcpa-software.com/}{FACSIMILE Kinetic Modelling
Software}; for historical reasons, the software and the format have been
widely used in conjunction with the MCM. The
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/extract.htt}{extraction tool} on
the MCM website can generate \texttt{.fac} files directly in FACSIMILE
format.

\subsection{FACSIMILE format} \label{subsec:facsimile-format}

Chemical reactions are described in FACSIMILE format using the following
notation:

\begin{verbatim}
% k : A + B = C + D ;
\end{verbatim}

where \texttt{k} is the rate coefficient, \texttt{A} and \texttt{B} are
the reactants, \texttt{C} and \texttt{D} are the products. The reaction
starts with the \texttt{\%} character and ends with the \texttt{;}
character. Comments - enclosed between the characters \texttt{*} and
\texttt{;} - can be added to the mechanism, and will be ignored by the
compiler. For example:

\begin{verbatim}
* conversion of A to C with rate coefficient of 1e-4 *;
% 1E-4 : A = C ;
\end{verbatim}

The mechanism file is processed by the script \texttt{tools/build.sh},
as explained in the {[}{[}model setup page\textbar{}2. Model Setup and
Execution{]}{]}. For the build process to work, the \texttt{.fac} file
must include four sections delimited by a single comment line, which
allows the script to recognize the beginning of each section: 1. Generic
rate coefficients 1. Complex reactions rate coefficients 1. Sum of
peroxy radicals (see below) 1. Chemical Reactions

These comment lines must always be present, even though the respective
sections can be empty. A minimal \texttt{.fac} file looks like this:

\begin{verbatim}
* Generic Rate Coefficients ;

* Complex reactions ;

* Peroxy radicals. ;

RO2 =  ;

* Reaction definitions. ;

% k : A + B = C ;
\end{verbatim}

A simple chemical mechanism in FACSIMILE format - with the first step of
the atmospheric oxidation of ethanol - is shown below, as an example.

\subsection{RO2 sum} \label{subsec:ro2-sum}

The sum of organic peroxy radicals (RO2) is a key component of the
Master Chemical Mechanism (see the MCM protocol papers:
\href{https://doi.org/10.1016/S1352-2310(96)00105-7}{Jenkin et al.,
Atmos. Environ., 31, 81-104, 1997} and
\href{https://doi.org/10.5194/acp-3-161-2003}{Saunders et al., Atmos.
Chem. Phys., 3, 161-180, 2003}). Since AtChem2 is designed primarily to
run models based upon the MCM, the \texttt{.fac} file must contain a
section with the RO2 sum. This section must be introduced by the comment
line \texttt{*\ Peroxy\ radicals.\ ;} (see above) and has the format:

\begin{verbatim}
RO2 = RO2a + RO2b + RO2c + ... ;
\end{verbatim}

where \texttt{RO2a}, \texttt{RO2b}, \texttt{RO2c}, are the organic
peroxy radicals in the chemical mechanism. If there are no organic
peroxy radicals in the mechanism (or if the mechanism is not based upon
the MCM), the RO2 sum must be left empty, e.g.:

\begin{verbatim}
RO2 = ;
\end{verbatim}

\emph{Important}: HO2 is a peroxy radical, but it is not an organic
molecule. Therefore it should NOT be included in the RO2 sum.

The RO2 sum is automatically generated from the mechanism file during
the build process, using the list of RO2 extracted from the MCM
database. AtChem2 includes the list of all the organic peroxy radicals
in version 3.3.1 of the MCM (\texttt{mcm/peroxy-radicals\_v3.3.1}),
which is used by default. Since v1.1, lists of organic peroxy radicals
from other versions of the MCM are also included in the \texttt{mcm/}
directory: see the file \texttt{mcm/INFO.md} for instructions on how to
use previous versions of the MCM with AtChem2.

\subsection{The MCM extraction tool} \label{subsec:the-mcm-extraction-tool}

The MCM website provides a convenient tool which can be used to download
the whole MCM, or subsets of it, in FACSIMILE format. After selecting
the species of interest in the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/roots.htt}{MCM browser}, add them
to the \emph{Mark List}, then proceed to the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/extract.htt}{MCM extraction tool}
and select \emph{FACSIMILE} as format. Make sure to tick the boxes:

\begin{verbatim}
[x] Include inorganic reactions?
[x] Include generic rate coefficients? FACSIMILE, FORTRAN and KPP formats only
\end{verbatim}

then press the \emph{Extract} button to download the generated
\texttt{.fac} file into a directory of choice (e.g., \texttt{model/};
see the {[}{[}model structure page\textbar{}1.2 Model Structure{]}{]}).
The mechanism can be modified with a text editor (if necessary) or
directly used in AtChem2. More information about the MCM browser and the
extractor tool can be found on the \href{http://mcm.leeds.ac.uk}{MCM
website}.

\subsection{The build process} \label{subsec:the-build-process}

Atchem2 uses a Python script (\texttt{tools/mech\_converter.py},
automatically called by \texttt{tools/build.sh} during the build
process) to convert the chemical mechanism into a format that can be
read by the Fortran code.

The script generates one Fortran file, one shared library, and four
configuration files from the \texttt{*.fac} file:

\begin{itemize}
\item
  \textbf{mechanism.f90} contains the equations, in Fortran code, to
  calculate the rate coefficients of each reaction. By default, it is
  placed in \texttt{model/configuration/}.
\item
  \textbf{mechanism.so} is the compiled version of
  \texttt{mechanism.f90}. By default, it is placed in
  \texttt{model/configuration/}.
\item
  \textbf{mechanism.species} contains the list of chemical species in
  the mechanism. By default, it is saved in
  \texttt{model/configuration/}. The file has no header. The first
  column is the \emph{ID number} of each species, the second column is
  the name of the species:
  \texttt{1\ O\ \ \ 2\ O3\ \ \ 3\ NO\ \ \ 4\ NO2}
\item
  \textbf{mechanism.reac} and \textbf{mechanism.prod} contain the
  reactants and the products (respectively) in each reaction of the
  mechanism. By default, it is saved in \texttt{model/configuration/}.
  The files have a 1 line header with the number of species, the number
  of reactions and the number of equations in the Generic Rate
  Coefficients and Complex Reactions sections. The first column is the
  \emph{ID number} of the reaction, the second column is the \emph{ID
  number} of the species (from \texttt{mechanism.species}) which are
  reactants/products in that reaction:
  \texttt{29\ 71\ 139\ numberOfSpecies\ numberOfReactions\ numberOfGenericComplex\ \ \ 1\ 1\ \ \ 2\ 1\ \ \ 3\ 1\ \ \ 3\ 2}
\item
  \textbf{mechanism.ro2} contains the organic peroxy radicals (RO2). By
  default, it is saved in \texttt{model/configuration/}. The file has a
  comment line header (Fortran style). The first column is the \emph{ID
  number} of the peroxy radical (from \texttt{mechanism.species}), the
  second column is the name of the peroxy radical as Fortran comment:
  \texttt{!\ Note\ that\ this\ file\ is\ generated\ by\ tools/mech\_converter.py\ based\ upon\ the\ file\ tools/mcm\_example.fac.\ Any\ manual\ edits\ to\ this\ file\ will\ be\ overwritten\ when\ calling\ tools/mech\_converter.py\ \ \ 23\ !CH3O2\ \ \ 26\ !C2H5O2\ \ \ 28\ !IC3H7O2\ \ \ 29\ !NC3H7O2}
\end{itemize}

The locations of the files generated during the build process can be
modified by changing the second and the third argument of the script
\texttt{tools/build.sh}. For more information and detailed instructions
go to: {[}{[}2. Model Setup and Execution{]}{]}.

\subsection{Example mechanism file} \label{subsec:example-mechanism-file}

\begin{verbatim}
* ------------------------------------------------------------------- *;
* SIMPLE CHEMICAL MECHANISM                                           *;
* Chemical mechanism for ethanol - from MCM v3.3.1                    *;
* ------------------------------------------------------------------- *;
*;
* Generic Rate Coefficients ;
*;
* Complex reactions ;
*;
* Peroxy radicals. ;
RO2 = HOCH2CH2O2 ;
*;
* Reaction definitions. ;
% 3.0D-12*EXP(20/TEMP)*0.05 : C2H5OH + OH = C2H5O ;
% 3.0D-12*EXP(20/TEMP)*0.9  : C2H5OH + OH = CH3CHO + HO2 ;
% 3.0D-12*EXP(20/TEMP)*0.05 : C2H5OH + OH = HOCH2CH2O2 ;
\end{verbatim}

% -------------------------------------------------------------------- %
\section{Model Parameters} \label{sec:parameters}

The \textbf{model parameters} are set in the text file
\texttt{model/configuration/model.parameters}; they control the general
setup of the model.

\begin{itemize}
\item
  \textbf{number of steps} and \textbf{step size}. The duration of the
  model run is determined by the number of steps and the step size (in
  seconds). The step size controls the frequency of the model output for
  the chemical species listed in \texttt{outputSpecies.config} (see
  {[}{[}2.6 Config Files{]}{]}), and for the environment variables, the
  photolysis rates, the diagnostic variables.\\
  For example, a model runtime of 2 hours, with output every 5 minutes,
  requires 24 steps with a step size of 300 seconds (24x300 = 7200 sec =
  2 hours). Possible values for these parameters are shown below, for
  reference.
\item
  \textbf{species interpolation method} and \textbf{conditions
  interpolation method}. Interpolation method used for the constrained
  chemical species, and for the constrained environment variables and
  the photolysis rates, respectively (see {[}{[}2.7 Constraints{]}{]}).
  Two interpolation methods are currently implemented in AtChem2:
  piecewise constant (\texttt{1}) and piecewise linear (\texttt{2}). The
  default option is \emph{piecewise linear interpolation}.
\item
  \textbf{rates output step size}. Frequency (in seconds) of the model
  output for the production and loss rates of selected species. The
  species for which this parameter is required are listed in
  \texttt{outputRates.config} (see {[}{[}2.6 Config Files{]}{]}).
\item
  \textbf{model start time}. Start time of the model (in seconds)
  calculated from midnight of the \textbf{day}, \textbf{month},
  \textbf{year} parameters (see below). For example, a start time of
  3600 means the model run starts at 1:00 in the morning and a start
  time of 43200 means the model run starts at midday. The \textbf{model
  stop time} is automatically calculated as:
  \texttt{model\ start\ time\ +\ (number\ of\ steps\ *\ step\ size))}.\\
  \emph{Important}: if one or more variables are constrained, the
  interval between the model start time and the model stop time must be
  equal or shorter than the time interval of the constrained data (see
  {[}{[}2.7 Constraints{]}{]}).
\item
  \textbf{jacobian output step size}. Frequency of the model output for
  the Jacobian matrix (in seconds). If the frequency is set to
  \texttt{0} (default option), the Jacobian matrix is not output. Note
  that the \texttt{jacobian.output} file generated by the model can be
  very large, especially if the chemical mechanism has many reactions
  and/or the model runtime is long.
\item
  \textbf{latitude} and \textbf{longitude}. Geographical coordinates (in
  degrees). By convention, latitude North is positive and latitude South
  is negative, longitude East is negative and longitude West is
  positive. Latitude and longitude are used only for the calculation of
  the Earth-Sun angles, which are needed for the MCM photolysis
  parameterisation (see {[}{[}2.5 Photolysis Rates and JFAC{]}{]}).
\item
  \textbf{day} and \textbf{month} and \textbf{year}. Start date of the
  model simulation. The model time is in seconds since midnight of the
  start date.
\item
  \textbf{reaction rates output step size}. Frequency (in seconds) of
  the model output for the reaction rates of every reaction in the
  chemical mechanism. The reaction rates are saved in the directory
  \texttt{model/output/reactionRates/} as one file for each model step,
  with the name of the file corresponding to the time in seconds. In
  previous versions of AtChem, this output was called
  \emph{instantaneous rates}.\\
  Note that this parameter is different from \textbf{rates output step
  size} (see above), which sets the frequency of a formatted output of
  reaction rates for selected species of interest. For more information
  go to: {[}{[}2.6 Config Files{]}{]}.
\end{itemize}

\subsection{Runtime reference values} \label{subsec:runtime-reference-values}

For 1 day at 15 minute intervals:

\begin{verbatim}
96      number of steps
900     step size
\end{verbatim}

For 2 days at 15 minute intervals:

\begin{verbatim}
192     number of steps
900     step size
\end{verbatim}

For 2 days at 1 minute intervals:

\begin{verbatim}
2880    number of steps
60      step size
\end{verbatim}

% -------------------------------------------------------------------- %
\section{Solver Parameters} \label{sec:solver}

The \textbf{solver parameters} are set in the text file
\texttt{model/configuration/solver.parameters}; they control the
behaviour of the ordinary differential equations (ODE) solver. A
complete explanation of these parameters can be found in the
\href{https://computation.llnl.gov/projects/sundials/sundials-software}{CVODE
documentation}.

\begin{itemize}
\item
  \textbf{atol} (positive real) and \textbf{rtol} (positive real):
  absolute and relative tolerance values for the solver. Standard values
  for these parameters are listed below, for reference.
\item
  \textbf{delta main} (positive real): linear convergence tolerance
  factor of the GMRES linear solver.
\item
  \textbf{lookback} (positive integer): maximum Krylov subspace
  dimension of the GMRES linear solver.
\item
  \textbf{maximum solver step size} (positive real): maximum size (in
  seconds) of the timesteps that the solver is allowed to use.
\item
  \textbf{maximum number of steps in solver} (positive integer): maximum
  number of steps used by the solver before reaching \textbf{tout}, the
  next output time.
\item
  \textbf{solver type} (integer): selects the linear solver to use:
  \texttt{1} for GMRES, \texttt{2} for GMRES preconditioned with a
  banded preconditioner, \texttt{3} for a dense solver. The default
  option is \texttt{2}.
\item
  \textbf{banded preconditioner upper bandwidth} (integer): used in the
  case that \texttt{solver\ type\ =\ 2}.
\item
  \textbf{banded preconditioner lower bandwidth} (integer): used in the
  case that \texttt{solver\ type\ =\ 2}.
\end{itemize}

\subsection{Solver reference values} \label{subsec:solver-reference-values}

Standard solver tolerance values:

\begin{verbatim}
1.0e-04     atol
1.0e-06     rtol
\end{verbatim}

% -------------------------------------------------------------------- %
\section{Environment Variables} \label{sec:envvar}

The \textbf{environment variables} define the physical parameters of the
box-model, such as temperature, pressure, humidity, latitude, longitude,
position of the sun, etc\ldots{} These variables are set in the text
file \texttt{model/configuration/environmentVariables.config}.

The environment variables can have a fixed (constant) value or can be
constrained to measured values (\texttt{CONSTRAINED}), in which case the
corresponding data file must be in the
\texttt{model/constraints/environment/} directory (see {[}{[}2.7
Constraints{]}{]}). Some environment variables can be calculated by the
model (\texttt{CALC}) and some can be deactivated if they are not used
by the model (\texttt{NOTUSED}).

By default, most environment variables are set to a fixed value,
corresponding to \emph{standard environmental conditions} (listed
below), or to \texttt{NOTUSED}.

\textbf{TEMP}

Ambient Temperature (K).

\begin{itemize}
\item
  fixed value
\item
  constrained
\end{itemize}

Default fixed value = 298.15

\textbf{PRESS}

Ambient Pressure (mbar).

\begin{itemize}
\item
  fixed value
\item
  constrained
\end{itemize}

Default fixed value = 1013.25

\textbf{RH}

Relative Humidity (\%). It is required only if \textbf{H2O} is set to
\texttt{CALC}, otherwise should be set to \texttt{NOTUSED}.

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  not used
\end{itemize}

Default = NOTUSED (-1)

\textbf{H2O}

Water Concentration (molecules cm-3).

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  calculated -\textgreater{} requires \textbf{RH} set to fixed value or
  \texttt{CONSTRAINED}
\end{itemize}

Default fixed value = 3.91e+17

\textbf{DEC}

Sun Declination (radians) is the angle between the center of the Sun and
Earth's equatorial plane.

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  calculated -\textgreater{} requires \textbf{DAY} and \textbf{MONTH},
  which are set in \texttt{model.parameters} (see {[}{[}2.2 Model
  Parameters{]}{]})
\end{itemize}

Default fixed value = 0.41

\textbf{BLHEIGHT}

Boundary Layer Height. It is required only if the model includes
emission or deposition processes (it must be used in the chemical
mechanism as a multiplier of the rate coefficient). The unit is
typically in cm, but it depends on how the processes are parameterized
in the chemical mechanism (see {[}{[}2.1 Chemical Mechanism{]}{]}).

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  not used
\end{itemize}

Default = NOTUSED (-1)

\textbf{DILUTE}

Dilution rate. It is required only if the model includes a dilution
process (it must be used in the chemical mechanism as a multiplier of
the rate coefficient). The unit is typically in s-1, but it depends on
how the process is parameterized in the chemical mechanism (see
{[}{[}2.1 Chemical Mechanism{]}{]}).

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  not used
\end{itemize}

Default value = NOTUSED (-1)

\textbf{JFAC}

Correction factor used to correct the photolysis rates (e.g., to account
for cloudiness). The calculated photolysis rates are scaled by JFAC,
which can have a value between \texttt{0} (photolysis rates go to zero)
and \texttt{1} (photolysis rates are not corrected). JFAC is NOT applied
to constant or constrained photolysis rates. For more information go to:
{[}{[}2.5 Photolysis Rates and JFAC{]}{]}.

\begin{itemize}
\item
  fixed value
\item
  constrained
\item
  calculated
\end{itemize}

Default fixed value = 1

\textbf{ROOF}

Flag to turn the photolysis rates ON/OFF. It is used in simulations of
environmental chamber experiments, where the roof of the chamber can be
opened/closed or the lights turned on/off.

When ROOF is set to \texttt{CLOSED} all the photolysis rates are zero,
including those that are constant or constrained; this is different than
setting JFAC to \texttt{0}, which only applies to the calculated
photolysis rates (see above). ROOF is the only environment variable that
cannot be set to \texttt{CONSTRAINED}.

Default value = OPEN

\subsection{Standard environmental conditions} \label{subsec:standard-environmental-conditions}

\begin{verbatim}
Temperature = 25C  
Pressure = 1 atm  
Relative Humidity = 50%  
Day, Month = 21 June
\end{verbatim}

% -------------------------------------------------------------------- %
\section{Photolysis rates} \label{sec:photolysis}

The photolysis rates are identified in {[}{[}FACSIMILE
format\textbar{}2.1 Chemical Mechanism{]}{]} as
\texttt{J\textless{}n\textgreater{}}, where \texttt{n} is an integer
determined by the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/parameters/photolysis.htt}{MCM
naming convention}. The photolysis rates are calculated by AtChem2 using
the
\href{http://mcm.leeds.ac.uk/MCM/parameters/photolysis_param.htt}{MCM
parametrization}, as explained in more detail below. Each photolysis
rate can also be set to a constant value or to constrained values.

The following rules apply:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  If a photolysis rate is set as constant, it assumes the given value.
  Any other photolysis rate, without an explicitly defined constant
  value, is set to zero.
\item
  If one or more photolysis rates are set to constrained (and none is
  set to constant), they assume the values given in the corresponding
  constraint files. Any other photolysis rate is calculated.
\item
  If no photolysis rate is set to constant or to constrained, the model
  calculates all the photolysis rates.
\end{enumerate}

The environment variable \texttt{ROOF} can also be used to turn the
photolysis rates ON/OFF, which is useful for simulations of some
environmental chamber experiments (see {[}{[}2.4 Environment
Variables{]}{]}).

\subsection{Constant photolysis rates} \label{subsec:constant-photolysis-rates}

The typical scenario for constant photolysis rates is the use of a lamp
in an environmental chamber. All the photolysis rates used in the
mechanism need to be given a value (in
\texttt{model/configuration/photolysisConstant.config}) otherwise they
will be set to zero. This approach allows the user to model individual
photolysis processes and/or to account for lamps that emit only in
certain spectral windows. The format of the
\texttt{photolysisConstant.config} file is described in the
{[}{[}configuration files page\textbar{}2.6 Config Files{]}{]}.

\subsection{Constrained photolysis rates} \label{subsec:constrained-photolysis-rates}

Photolysis rates can be constrained to measured values. In this case,
the name of the constrained photolysis rate (e.g., \texttt{J2}) must be
in \texttt{model/configuration/photolysisConstrained.config} and a file
with the constraint data must be present in
\texttt{model/constraints/photolysis/}. For more information go to:
{[}{[}2.6 Config Files{]}{]} and {[}{[}2.7 Constraints{]}{]}.

It is not always possibile to measure - and therefore constrain - all
the required photolysis rates. The photolysis rates that are not
constrained (i.e., not listed in \texttt{photolysisConstrained.config})
are calculated using the MCM parametrization.

\subsection{Calculated photolysis rates} \label{subsec:calculated-photolysis-rates}

AtChem2 implements the parametrization of photolysis rates used by the
Master Chemical Mechanism. It is described in the MCM protocol papers:
\href{https://doi.org/10.1016/S1352-2310(96)00105-7}{Jenkin et al.,
Atmos. Environ., 31, 81, 1997} and
\href{https://doi.org/10.5194/acp-3-161-2003}{Saunders et al., Atmos.
Chem. Phys., 3, 161, 2003}.

The MCM parametrization calculates the photolysis rate of a reaction
(\texttt{J}) with the equation:

\begin{verbatim}
J = l * (cosX)^m * exp(-n * secX) * tau
\end{verbatim}

where \texttt{l}, \texttt{m}, \texttt{n} are empirical parameters,
\texttt{cosX} is the cosine of the solar zenith angle, \texttt{secX} is
the inverse of \texttt{cosX} (i.e., \texttt{secX\ =\ 1/cosX}) and
\texttt{tau} is the transmission factor. The empirical parameters are
different for each version of the MCM. AtChem2 v1.1 includes the
empircal parameters for
\href{http://mcm.leeds.ac.uk/MCM/parameters/photolysis_param.htt}{version
3.3.1} in the file \texttt{mcm/photolysis-rates\_v3.3.1}. This file also
contains the transmission factor \texttt{tau}, which can be changed by
the user (by default \texttt{tau\ =\ 1}). It is possible to use previous
versions of the MCM parametrization: see the file \texttt{mcm/INFO.md}
for instructions.

The solar zenith angle is calculated by AtChem2 using latitude,
longitude, time of the day and sun declination (see {[}{[}2.2 Model
Parameters{]}{]} and {[}{[}2.4 Environment Variables{]}{]}). The
calculation is detailed in ``The Atmosphere and UV-B Radiation at Ground
Level'' (\href{https://doi.org//10.1007/978-1-4899-2406-3_1}{S.
Madronich, Environmental UV Photobiology, 1993}).

\subsection{JFAC} \label{subsec:jfac}

Measurements of ambient photolysis rates typically show short-term
variability, due to the changing meteorological conditions (clouds,
rain, etc\ldots{}). This information is retained in the constrained
photolysis rates, but it is lost in the calculated ones. To account for
this, the calculated photolysis rates can be scaled by a correction
factor (\texttt{JFAC}), as explained below.

The environment variable \texttt{JFAC} is a constant or time-dependent
parameter that can be used to correct the calculated photolysis rates
for external factors not taken into account by the MCM parametrization,
such as cloudiness. \texttt{JFAC} is defined as the ratio between a
measured and the calculated photolysis rate. Typically \texttt{J4} (the
photolysis rate of NO2) is used for this purpose, as it is one of the
most frequently measured photolysis rates.

\begin{verbatim}
JFAC = j(NO2)/J4
\end{verbatim}

where \texttt{j(NO2)} is the measured value and \texttt{J4} is
calculated with the MCM parametrization (see above). \texttt{JFAC} is by
default 1, meaning that the calculated photolyis rates are not scaled;
it can be set to any value between 0 and 1 (see {[}{[}2.4 Environment
Variables{]}{]}) or it can be constrained (see {[}{[}2.7
Constraints{]}{]}). Note that only the photolysis rates calculated with
the MCM parameterization are scaled by \texttt{JFAC}, the constrained
and the constant photolysis rates are not.

\texttt{JFAC} can also be calculated at runtime. To do so, \texttt{JFAC}
should be set to the name of the photolysis rate to be used as reference
(e.g., \texttt{J4}) in
\texttt{model/configuration/environmentVariables.config}. There should
be an associated constraint file in
\texttt{model/constraints/environment/}. \textbf{Important}: this option
is not working very well in the current version of AtChem2, so it is
suggested to calculate \texttt{JFAC} offline and to constrain it (see
issue \href{https://github.com/AtChem/AtChem2/issues/16}{\#16}).

% -------------------------------------------------------------------- %
\section{Config Files} \label{sec:config}

The \textbf{configuration files} contain the settings for the initial
conditions, the constraints and the output of the model. These files
complement the configuration settings of the model (in
\texttt{model.parameters}) and of the solver (in
\texttt{solver.parameters}), which are in the same directory. For more
information go to: {[}{[}2.2 Model Parameters{]}{]} and {[}{[}2.3 Solver
Parameters{]}{]}).

The configuration files have the extension \texttt{.config} and, by
default, are in the directory \texttt{model/configuration/}. This
directory also contains the files generated during the {[}{[}build
process\textbar{}2. Model Setup and Execution{]}{]} which describe the
chemical mechanism (\texttt{mechanism.species}, \texttt{mechanism.reac},
\texttt{mechanism.prod}, \texttt{mechanism.ro2}), as explained in the
{[}{[}chemical mechanism page\textbar{}2.1 Chemical Mechanism{]}{]}. The
location of the configuration files can be modified by changing the
arguments of the script \texttt{tools/build.sh} (see {[}{[}2. Model
Setup and Execution{]}{]}).

The content and the format of the \texttt{.config} files are described
below. Note that the names of some files have changed with the release
of \textbf{version 1.1} (November 2018).

\subsection{environmentVariables.config} \label{subsec:environmentvariables}

This file contains the settings for the environment variables, which are
described in detail in the related {[}{[}wiki page\textbar{}2.4
Environment Variables{]}{]}. If an environment variable is constrained,
there must be a corresponding data file in
\texttt{model/constraints/environment/} (see {[}{[}2.7
Constraints{]}{]}).

\subsection{initialConcentrations.config} \label{subsec:initialconcentrations}

This file contains the initial concentrations of the chemical species.
The first column is the list of initialized species, the second column
is the corresponding concentration at \texttt{t\ =\ 0} (in
\textbf{molecules cm-3}). For example:

\begin{verbatim}
NO      378473308.14
NO2     86893908168.9
O3      1.213e+12
CH4     4.938e+13
\end{verbatim}

The chemical species not included in this file are automatically
initialized to the default value \texttt{0}. It is not necessary to
initialize the constant and the constrained species (i.e., those listed
in \texttt{speciesConstant.config} and
\texttt{speciesConstrained.config}).

The environment variables are set in
\texttt{environmentVariables.config} (see above) and should not be
included in this file.

\subsection{outputRates.config} \label{subsec:outputrates}

This file (called \texttt{productionRatesOutput.config} and
\texttt{lossRatesOutput.config} in v1.0) lists the chemical species for
which detailed production rates and loss rates are required. The file
has one column, with one species per line.

The frequency of this output is controlled by the \textbf{rates output
step size} parameter in \texttt{model.parameters} (see {[}{[}2.2 Model
Parameters{]}{]}). The format of the corresponding output files -
\texttt{lossRates.output} and \texttt{productionRates.output} - is
designed to facilitate the analysis of production and destruction rates
for selected species of interests (rather than processing the output
files in \texttt{model/output/reactionRates/}):

\begin{verbatim}
     time        speciesNumber    speciesName    reactionNumber         rate        reaction

3.600000E+003           8               OH              15         0.000000E+000    O1D=OH+OH
3.600000E+003           8               OH              20         0.000000E+000    HO2+O3=OH
3.600000E+003           9               HO2             16         0.000000E+000    OH+O3=HO2
3.600000E+003           9               HO2             17         0.000000E+000    OH+H2=HO2

7.200000E+003           8               OH              15         0.000000E+000    O1D=OH+OH
7.200000E+003           8               OH              20         0.000000E+000    HO2+O3=OH
7.200000E+003           9               HO2             16         0.000000E+000    OH+O3=HO2
7.200000E+003           9               HO2             17         0.000000E+000    OH+H2=HO2
\end{verbatim}

\subsection{outputSpecies.config} \label{subsec:outputspecies}

This file (called \texttt{concentrationOutput.config} in v1.0) lists the
chemical species for which the model output is required. The current
version of AtChem2 limits the number of species that can be output to
100, although the user can modify the Fortran code to increase this
number. The file has one column, with one species per line.

The frequency of this output is controlled by the \textbf{step size}
parameter in \texttt{model.parameters} (see {[}{[}2.2 Model
Parameters{]}{]}).

\subsection{photolysisConstant.config} \label{subsec:photolysisconstant}

This file lists the photolysis rates that are constant. The file has
three columns: the first column is the number that identifies the
photolysis rate (e.g., \texttt{1}), the second column is the value of
the photolysis rate in \textbf{s-1} (e.g., \texttt{1e-5}), the third
column is the name of the photolysis rate (e.g., \texttt{J1}). The
photolysis rates are named according to the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/parameters/photolysis.htt}{MCM
naming convention}. If no photolysis rate is constant, the file should
be left empty.

If one or more photolysis rates is set to a constant value, the others
(i.e., those not listed in \texttt{photolysisConstants.config}) are set
to zero. For more information go to: {[}{[}2.5 Photolysis Rates and
JFAC{]}{]}.

\subsection{photolysisConstrained.config} \label{subsec:photolysisconstrained}

This file (called \texttt{constrainedPhotoRates.config} in v1.0) lists
the photolysis rates that are constrained. The file has one column, with
one photolysis rate per line (e.g., \texttt{J1}). The photolysis rates
are named according to the
\href{http://mcm.leeds.ac.uk/MCMv3.3.1/parameters/photolysis.htt}{MCM
naming convention}. If no photolysis rate is constrained, the file
should be left empty. If a photolysis rate is constrained, there must be
a corresponding data file in \texttt{model/constraints/photolysis/} (see
{[}{[}2.7 Constraints{]}{]}).

The photolysis rates that are not listed in
\texttt{photolysisConstrained.config} are calculated by AtChem2 using
the MCM parametrization and the parameters in
\texttt{mcm/photolysis-rates\_v3.3.1}. Older versions of the MCM
photolysis parametrization can be used, as explained in the file
\texttt{mcm/INFO.md}. For more information go to: {[}{[}2.5 Photolysis
Rates and JFAC{]}{]}.

\subsection{speciesConstant.config} \label{subsec:speciesconstant}

This file (called \texttt{constrainedFixedSpecies.config} in v1.0) lists
the chemical species that are constant. The file has two columns: the
first column is the list of constant species, the second column is the
corresponding concentration (in \textbf{molecules cm-3}). If no chemical
species is constant, the file should be left empty.

If a chemical species is constant, it does not need to be initialized:
the values set in \texttt{speciesConstant.config} override those set in
\texttt{initialConcentrations.config}.

\subsection{speciesConstrained.config} \label{subsec:speciesconstrained}

This file (called \texttt{constrainedSpecies.config} in v1.0) lists the
chemical species that are constrained. The file has one column, with one
species per line. If no chemical species is constrained, the file should
be left empty. If a chemical species is constrained, there must be a
corresponding data file in \texttt{model/constraints/species/} (see
{[}{[}2.7 Constraints{]}{]}).

If a chemical species constrained, it does not need to be initialized:
the values set in \texttt{speciesConstrained.config} override those set
in \texttt{initialConcentrations.config}.

% -------------------------------------------------------------------- %
\section{Constraints} \label{sec:constraints}

AtChem2 can be run in two modes:

\begin{itemize}
\item
  unconstrained: all variables are calculated by the model from the
  initial conditions, set in the {[}{[}model configuration
  files\textbar{}2.6 Config Files{]}{]}.
\item
  constrained: one or more variables are constrained, i.e.~the solver
  forces their value to a given value. The variables that are not
  constrained are calculated by the model.
\end{itemize}

The constrained values must be provided as separate files for each
constrained variable. The format of the constraint files is described
below. By default, the files with the constraining data are in
\texttt{model/constraints/species/} for the chemical species,
\texttt{model/constraints/environment/} for the environment variables,
and \texttt{model/constraints/photolysis/} for the photolysis rates. The
default directories can be modified by changing the arguments of the
\texttt{atchem2} executable (see {[}{[}2. Model Setup and
Execution{]}{]}).

\subsection{Constrained variables} \label{subsec:constrained-variables}

\subsubsection{Environment variables} \label{environment-variables}

All environment variables, except \texttt{ROOF}, can be constrained. To
do so, set the variable to \texttt{CONSTRAINED} in
\texttt{model/configuration/environmentVariables.config} and create the
file with the constraining data. The name of the file must be the same
as the name of the variable, e.g.~\texttt{TEMP} (without extension). See
also: {[}{[}2.4 Environment Variables{]}{]}.

\subsubsection{Chemical species} \label{chemical-species}

Any chemical species in the chemical mechanism can be constrained. To do
so, add the name of the species to
\texttt{model/configuration/speciesConstrained.config} and create the
file with the constraining data. The name of the file must be the same
as the name of the chemical species, e.g.~\texttt{CH3OH} (without
extension). See also: {[}{[}2.6 Config Files{]}{]}.

\subsubsection{Photolysis rates} \label{photolysis-rates}

Any of the photolysis rates in the chemical mechanism can be
constrained. The photolysis rates are identified as
\texttt{J\textless{}n\textgreater{}}, where \texttt{n} is an integer
(see {[}{[}2.5 Photolysis Rates and JFAC{]}{]}). To constrain a
photolysis rate add its name (\texttt{Jn}) to
\texttt{model/configuration/photolysisConstrained.config} and create the
file with the constraining data. The name of the file must be the same
as the name of the photolysis rate, e.g.~\texttt{J4} (without
extension). See also {[}{[}2.6 Config Files{]}{]}.

\subsection{Constraint files} \label{subsec:constraint-files}

The files with the constraining data are text files with two columns
separated by spaces. The first column is the time in \textbf{seconds}
from midnight of day/month/year (see {[}{[}2.2 Model Parameters{]}{]}),
the second column is the value of the variable in the appropriate unit.
For the chemical species the unit is \textbf{molecules cm-3} and for the
photolysis rates the unit is \textbf{s-1}; for the environment variables
see the related {[}{[}wiki page\textbar{}2.4 Environment
Variables{]}{]}. For example:

\begin{verbatim}
-900   73.21
0      74.393
900    72.973
1800   72.63
2700   72.73
3600   69.326
4500   65.822
5400   63.83
6300   64.852
7200   64.739
\end{verbatim}

The time in the first column of a constraint file can be negative.
AtChem2 interprets the negative timestamps as ``seconds \emph{before}
midnight of day/month/year'' (see {[}{[}2.2 Model Parameters{]}{]}).
This can be useful to allow correct interpolation of the variables at
the beginning of the model run (see below).

\textbf{Important.} The constraints must cover the same amount of time,
or preferably more, as the intended model runtime. For example: if the
model starts at 42300 seconds and stops at 216000 seconds, the first and
the last data points in a constraint file must have a timestamp of 42300
(or lower) and 21600 (or higher), respectively.

\subsection{Interpolation} \label{subsec:interpolation}

Constraints can be provided at different timescales. Typically, the
constraining data come from direct measurements and it is a very common
for different instruments to sample at different frequencies. For
example, ozone and nitrogen oxides can be measured once every minute,
but most organic compounds can be measured only once every hour.

The user can average the constraints so that they are all at the same
timescale or can use the data with the original timestamps. Both
approaches have advantages and disadvantages in terms of how much
pre-processing work is required, and in terms of model accuracy and
integration speed. Whether all the constraints have the same timescale
or not, the solver interpolates between data points using the
interpolation method selected in the
\texttt{model/configuration/model.parameters} file (see {[}{[}2.2 Model
Parameters{]}{]}). The default interpolation method is piecewise linear,
but piecewise constant interpolation is also available.

The photolysis rates and the environment variables are evaluated by the
solver when needed - each is interpolated individually, only when
constrained. This happens each time the function
\texttt{mechanism\_rates()} is called from \texttt{FCVFUN()}, and
therefore is controlled by \textbf{CVODE} as it completes the
integration. In a similar way, the interpolation routine for the
chemical species is called once for each of the constrained species in
\texttt{FCVFUN()}, plus once when setting the initial conditions of each
of the constrained species.

As mentioned above, the model start and stop time \emph{must be} within
the time interval of the constrained data to avoid interpolation errors
or model crash. If data is not supplied for the full runtime interval,
then the \emph{final} value will be used for all times both \emph{before
the first data point} and \emph{after the last data point}. This
behaviour is likely to change in future versions, at least to avoid the
situation where the last value is used for all times before the first
(see issue \href{https://github.com/AtChem/AtChem2/issues/294}{\#294}).

A warning is printed for all evaluations outside of the supplied time
interval. Users may find it useful to supply data that covers a short
time \emph{beyond} the final model time, which may be used by the
solver.

% -------------------------------------------------------------------- %
\section{Tools} \label{sec:tools}

The \texttt{tools/} directory contains a number of auxiliary scripts to
install, build and compile AtChem2, and to plot the results of the
model:

\begin{itemize}
\item
  shell script to compile the model: \texttt{build.sh}.
\item
  Python scripts to process the chemical mechanism:
  \texttt{fix\_mechanism\_fac.py}, \texttt{mech\_converter.py}.
\item
  Python scripts to enforce a consistent {[}{[}coding style\textbar{}3.2
  Style Guide{]}{]}: \texttt{fix\_indent.py}, \texttt{fix\_style.py}.
\item
  Ruby script to run the unit tests: \texttt{fruit\_generator.rb}.
\item
  example chemical mechanism in FACSIMILE format:
  \texttt{mcm\_example.fac}.
\item
  \texttt{install/} directory containing scripts to install the
  {[}{[}dependencies\textbar{}1.1 Dependencies{]}{]}.
\item
  \texttt{plot/} directory containing scripts to plot the model results.
\end{itemize}

In addition, the \texttt{tools/} directory contains a copy of the
\texttt{Makefile}, which has to be copied to the \emph{main directory}
and modified as explained in the {[}{[}installation page\textbar{}1.
Installation{]}{]}.

\subsection{Plot tools} \label{subsec:plot-tools}

The plotting scripts in \texttt{tools/plot/} are only intended to give a
quick view of the model results. It is suggested to use a proper data
analysis software (e.g., R, Octave/MATLAB, Igor, Origin, etc\ldots{}) to
process and analyze the model results. The scripts are written in
various programming languages, but they all produce the same output: a
file called \texttt{atchem2\_output.pdf} in the given directory (e.g.,
\texttt{model/output/}).

From the \emph{main directory}:

\begin{verbatim}
gnuplot -c tools/plot/plot-atchem2.gp model/output/
octave tools/plot/plot-atchem2.m model/output/
python tools/plot/plot-atchem2.py model/output/
Rscript --vanilla tools/plot/plot-atchem2.r model/output/
\end{verbatim}

\emph{N.B.}: the matlab script (\texttt{plot-atchem2.m}) is compatible
with both Octave and MATLAB. GNU Octave is an open-source implementation
of MATLAB.
