\chapter{Solver Parameters}

The \textbf{solver parameters} are set in the text file
\texttt{model/configuration/solver.parameters}; they control the
behaviour of the ordinary differential equations (ODE) solver. A
complete explanation of these parameters can be found in the
\href{https://computation.llnl.gov/projects/sundials/sundials-software}{CVODE
documentation}.

\begin{itemize}
\item
  \textbf{atol} (positive real) and \textbf{rtol} (positive real):
  absolute and relative tolerance values for the solver. Standard values
  for these parameters are listed below, for reference.
\item
  \textbf{delta main} (positive real): linear convergence tolerance
  factor of the GMRES linear solver.
\item
  \textbf{lookback} (positive integer): maximum Krylov subspace
  dimension of the GMRES linear solver.
\item
  \textbf{maximum solver step size} (positive real): maximum size (in
  seconds) of the timesteps that the solver is allowed to use.
\item
  \textbf{maximum number of steps in solver} (positive integer): maximum
  number of steps used by the solver before reaching \textbf{tout}, the
  next output time.
\item
  \textbf{solver type} (integer): selects the linear solver to use:
  \texttt{1} for GMRES, \texttt{2} for GMRES preconditioned with a
  banded preconditioner, \texttt{3} for a dense solver. The default
  option is \texttt{2}.
\item
  \textbf{banded preconditioner upper bandwidth} (integer): used in the
  case that \texttt{solver\ type\ =\ 2}.
\item
  \textbf{banded preconditioner lower bandwidth} (integer): used in the
  case that \texttt{solver\ type\ =\ 2}.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{solver-reference-values}{%
\subsection{Solver reference values}\label{solver-reference-values}}

Standard solver tolerance values:

\begin{verbatim}
1.0e-04     atol
1.0e-06     rtol
\end{verbatim}
